Utilizing Parallel Workers: \\LLVM's Vectorization Plan

### 2024-11-18 paper/outline.tex
add outline as seperate document

\begin{enumerate}
  \item Modern CPUs (SIMD-Registers, SIMD ISAs)
  \item SIMD performance gains
  \item Auto-Vectorization to help programmers better utilize SIMD functionality out of the box
  \item LLVM as popular, platform-agnostic compiler framework
  \item auto-vectorization complexity $\rightarrow$ wish for well designed, 
  interoperable system (VPlan) $\rightarrow$ ongoing 8 year long effort started by Intel
  \item rough paper outline
\end{enumerate}

@@

\section{Background}
\label{sec:background}
\begin{enumerate}
  \item Vectorization explanation (with Code\hyp{}Examples)
  \begin{itemize}
    \item Loop-Level Vectorization (Inner / Outer-Loops~\cite{llvmextloopvec,intelouterloop,10.5555/2337159.2337210,llvmouterloop,10.1145/1454115.1454119})
    \item Function Vectorization~\cite{llvmextloopvec}
    \item Superword-Level Parallelism\\Vectorization~\cite{10.1145/349299.349320}~(maybe already \cite{10.1145/3519939.3523701})
  \end{itemize}
  \item Vectorization Drawbacks~\cite{llvmveccontrolflow,intelvecessen} (reuse code\hyp{}examples from above)
  \begin{itemize}
    \item not always legal / applicable, e.g.: backwards data-dependencies, pointer aliasing, 
    platform backwards-compatability / support, control flow complexity (function calls, register pressure), FP inaccuracies (\texttt{-ffast-math})
    \item possible performance loss, e.g.: costly conversions (horizontal aggregations, integer division) resulting in slower overall code
    on average
    \item larger code size, e.g.: need for scalar loop-tail / epilogue
    \item (potential security vulnerablities~\cite{2023arXiv230201131K})
  \end{itemize}
  \item OpenMP (very short)
  \item Auto-Vectorization in Compilers (short)
  \item LLVM (very brief introduction)~\cite{10.5555/977395.977673}
  \item LLVM 'recent' transition from Loop Vectorizer + SLP Vectorizer to 
  Vectorization Plan (ongoing 8 year long refactoring and improvment 
  effort started by Intel)
\end{enumerate}

>>>>

\begin{enumerate}
  \item Modern CPUs (SIMD-Registers, SIMD ISAs)
  \item SIMD performance gains
  \item Auto-Vectorization to help programmers utilize SIMD functionality out of the box 
  (explicit vectorization as time consuming especially when supporting multiple architectures)
  \item LLVM as popular, platform-agnostic compiler framework~\cite{10.5555/977395.977673}
  \item auto-vectorization complexity $\rightarrow$ wish for well designed, 
  interoperable system (VPlan) $\rightarrow$ transition from Loop Vectorizer + SLP Vectorizer to 
  Vectorization Plan $\rightarrow$ ongoing 8 year long effort started by Intel
  \item rough paper outline
\end{enumerate}

@@

\section{Background}
\label{sec:background}
\begin{enumerate}
  \item Vectorization explanation (with Code\hyp{}Examples)
  \begin{itemize}
    \item Loop-Level Vectorization (Inner / Outer-Loops~\cite{llvmextloopvec,intelouterloop,10.5555/2337159.2337210,llvmouterloop,10.1145/1454115.1454119})
    \item Function Vectorization~\cite{llvmextloopvec}
    \item Superword-Level Parallelism\\Vectorization~\cite{10.1145/349299.349320}~(maybe already \cite{10.1145/3519939.3523701})
  \end{itemize}
  \item Vectorization Constraints~\cite{llvmveccontrolflow,intelvecessen} (reuse code\hyp{}examples from above)
  \begin{itemize}
    \item not always legal / applicable, e.g.: backwards data-dependencies, pointer aliasing, 
    platform backwards-compatability / support, control flow complexity (function calls, register pressure), FP inaccuracies (\texttt{-ffast-math})
    \item possible performance loss, e.g.: costly conversions (horizontal aggregations, integer division) resulting in slower overall code
    on average
    \item larger code size, e.g.: need for scalar loop-tail / epilogue
    \item (potential security vulnerablities~\cite{2023arXiv230201131K})
  \end{itemize}
\end{enumerate}

### 2024-11-18 paper/paper.tex
add outline as seperate document

\begin{abstract}
Your Abstract.
\end{abstract}

@@

\section{Introduction}
\label{sec:introduction}

\begin{enumerate}
  \item 
\end{enumerate}

@@

\section{Background}
\label{sec:background}
\begin{enumerate}
  \item (SIMD explanation necessary?)
  \item Vectorization explanation (with Code-Examples?)
  \begin{enumerate}
    \item Inner-Loop
    \item Outer-Loop
    \item Superword Level Parallelism~\cite{10.1145/349299.349320}~(maybe already \cite{10.1145/3519939.3523701})
  \end{enumerate}
  \item Vectorization Drawbacks (reuse code-examples from above)
  \begin{itemize}
    \item possible performance loss
    \item larger code size
    \item (potential security vulnerablities~\cite{2023arXiv230201131K})
  \end{itemize}
  \item LLVM short introduction
\end{enumerate}

@@

\section{LLVM's Vectorization Plan}
\label{sec:vplan}
\begin{enumerate}
  \item Cost-Model (related work?~\cite{POHL2020102106})
\end{enumerate}

@@

\section{Related Work}
\label{sec:relatedwork}
\begin{enumerate}
  \item Auto-Vectorization in GCC~\cite{autovecoptgcc, vecoptgcc}
  \item Comparison with GCC~\cite{gccllvmveccomp,10.1155/2021/3264624}
  \item (LLM-based Vectorization~\cite{taneja2024llmvectorizerllmbasedverifiedloop}, also if included maybe better fitting in "Future Work")
\end{enumerate}

@@

\section{Summary and Future Work}
\label{sec:summary}
\begin{enumerate}
  \item VPlan Summary and Future Roadmap~\cite{llvmvplan}
  \item Comparison/Outlook auto-vectorization vs. manual vectorization~\cite{10.1155/2021/3264624,9802745}
\end{enumerate}

>>>>

\begin{abstract}
Lorem ipsum
\end{abstract}

@@

\section{Introduction}
\label{sec:introduction}
Lorem ipsum

@@

\section{Background}
\label{sec:background}
Lorem ipsum

@@

\section{LLVM's Vectorization Plan}
\label{sec:vplan}
Lorem ipsum

@@

\section{Related Work}
\label{sec:relatedwork}
Lorem ipsum

@@

\section{Summary and Future Work}
\label{sec:summary}
Lorem ipsum
